I can see clearly now: reinterpreting statistical significance

### 2018-10-09 doc.tex

Using the language of ``statistical clarity'' would help researchers escape the trap of accepting the null hypothesis, reduce multiple types of misinterpretation associated with NHST, and allow for the continued use of NHST to determine whether we can be confident about the sign of a difference (e.g. see \citealt{Abelson1997} for arguments for NHST \bmb{what's in Abelson? How much of our argument do they anticipate?}). The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the \pval is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant. ``Clarity,'' on the other hand, is an apt term for what NHST actually evaluates. \citet{jones2000sensible} and \citet{robinson2001past} suggest that researchers should report $p < 0.05$ using language such as ``the direction of the differences among the treatments was undetermined''. While this is a step in the right direction, replacing ``significance'' with ``clarity'' is a more complete, direct, and universal change that will improve communication and reduce sloppy thinking.

@@

This correlation was significant in males ($\rho=0.35$, P \textless 0.05) but not females ($\rho=0.35$, NS). \ldots \ourcomment{The authors later write as though they have demonstrated a difference between males and females}
& Although males and females show the same correlation coefficient ($\rho=0.35$), the sign of the coefficient is statistically clear only in males  \ldots \ourcomment{This phrasing may suggest to the authors that confidence intervals are called for.}
\\
\rowcolor{lightgray}
\ldots risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but no association was found between HCV coinfection and low BMD in men 
(adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots  \ourcomment{The authors do not directly compare men and women, but they state:}
The precise mechanisms for the association between viral hepatitis and low BMD in HIV-infected women but not men 
remain unclear.
& 
\ldots  risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but the association between HCV coinfection and low BMD in men was not
statistically clear (adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots  Pursuing biological differences between women and men in the effect of HIV on BMD would be premature given these results.
\\
\end{tabular}
\captionof{table}{Examples of misinformed language in peer-reviewed papers (citations withheld out of courtesy, available by request) using our proposed  language of clarity. \bmb{do we want to comment on our search strategy? Doesn't need to be formal [this isn't a meta-analysis], but might be of interest to readers}}
\label{quotetab}
\end{table}

@@

\bmb{this section seems a bit ``grab-baggy''. Can we restructure it so it feels more coherent? I don't mind including these points, but I would rather leave them out than dilute the message/bog the reader down. Do we want to mention the more difficult case of NHST on non-negative domains (i.e. joint tests of multiple parameters, tests of variances, etc., where we can't interpret \pval as clarity about signs?)}

>>>>

Using the language of ``statistical clarity'' would help researchers escape the trap of accepting the null hypothesis, reduce multiple types of misinterpretation associated with accepting the null hypothesis, and allow for the continued use of NHST as a method of inference for for whether we can be confident of the sign of a difference (e.g. see \citealt{Abelson1997} for arguments for NHST \bmb{what's in Abelson? How much of our argument do they anticipate?}). The use of ``significance'' to describe the results of hypothesis tests is deeply, and sometimes subtly, misleading, because it is at odds with other meanings of the word: the \pval is not an accurate gauge of whether a result is large in magnitude, biologically important, or relevant. ``Clarity,'' on the other hand, is an apt term for what NHST actually evaluates. \citet{jones2000sensible} and \citet{robinson2001past} suggest that researchers should report $p < 0.05$ using language such as ``the direction of the differences among the treatments was undetermined''. While this is a step in the right direction, replacing ``significance'' with ``clarity'' is a more complete, direct, and universal change that will improve communication and reduce sloppy thinking.

@@

This correlation was significant in males ($\rho=0.35$, P \textless 0.05) but not females ($\rho=0.35$, NS). \ldots \ourcomment{The authors later write as though they have demonstrated a difference between males and females}
& Although males and females show the same correlation coefficient ($\rho=0.35$), the sign of the coefficient is statistically clear only in males  \ldots \ourcomment{This phrasing may suggest to the authors that confidence intervals are called for.}
\\
\rowcolor{lightgray}
\ldots risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but no association was found between HCV coinfection and low BMD in men 
(adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots  \ourcomment{The authors do not directly compare men and women, but they state:}
The precise mechanisms for the association between viral hepatitis and low BMD in HIV-infected women but not men 
remain unclear.
& 
\ldots  risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but the association between HCV coinfection and low BMD in men was not
statistically clear (adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots  Pursuing biological differences between women and men in the effect of HIV on BMD would be premature given these results.
\\
\end{tabular}
\captionof{table}{Examples of misinformed language in peer-reviewed papers (citations withheld out of courtesy, available by request) using our proposed  language of clarity.}
\label{quotetab}
\end{table}

@@

\bmb{this section seems a bit ``grab-baggy''. Can we restructure it so it feels more coherent? I don't mind including these points, but I would rather leave them out than dilute the message/bog the reader down.}
